### 2016-06-15 presentacion/paper.tex
arrancamos con la presentacion del paper

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Beamer Presentation
% LaTeX Template
% Version 1.0 (10/11/12)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND THEMES
%----------------------------------------------------------------------------------------

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\mode<presentation> {

% The Beamer class comes with a number of default slide themes
% which change the colors and layouts of slides. Below this is a list
% of all the themes, uncomment each in turn to see what they look like.

%\usetheme{default}
%\usetheme{AnnArbor}
%\usetheme{Antibes}
%\usetheme{Bergen}
%\usetheme{Berkeley}
%\usetheme{Berlin}
%\usetheme{Boadilla}
%\usetheme{CambridgeUS}
%\usetheme{Copenhagen}
%\usetheme{Darmstadt}
%\usetheme{Dresden}
%\usetheme{Frankfurt}
%\usetheme{Goettingen}
%\usetheme{Hannover}
%\usetheme{Ilmenau}
%\usetheme{JuanLesPins}
%\usetheme{Luebeck}
\usetheme{Madrid}
%\usetheme{Malmoe}
%\usetheme{Marburg}
%\usetheme{Montpellier}
%\usetheme{PaloAlto}
%\usetheme{Pittsburgh}
%\usetheme{Rochester}
%\usetheme{Singapore}
%\usetheme{Szeged}
%\usetheme{Warsaw}

% As well as themes, the Beamer class has a number of color themes
% for any slide theme. Uncomment each of these in turn to see how it
% changes the colors of your current slide theme.

%\usecolortheme{albatross}
%\usecolortheme{beaver}
%\usecolortheme{beetle}
%\usecolortheme{crane}
%\usecolortheme{dolphin}
%\usecolortheme{dove}
%\usecolortheme{fly}
%\usecolortheme{lily}
%\usecolortheme{orchid}
%\usecolortheme{rose}
%\usecolortheme{seagull}
%\usecolortheme{seahorse}
%\usecolortheme{whale}
%\usecolortheme{wolverine}

%\setbeamertemplate{footline} % To remove the footer line in all slides uncomment this line
%\setbeamertemplate{footline}[page number] % To replace the footer line in all slides with a simple slide count uncomment this line

%\setbeamertemplate{navigation symbols}{} % To remove the navigation symbols from the bottom of all slides uncomment this line
}

\usepackage{graphicx} % Allows including images
\usepackage{booktabs} % Allows the use of \toprule, \midrule and \bottomrule in tables

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\title[No Silver Bullet]{Frederick P. Brooks \\ No Silver Bullet \\ Essence and Accident in Software Engineering} % The short title appears at the bottom of every slide, the full title is only on the title page

\author{Grupo 3} % Your name
\institute[UBA - ISW2] % Your institution as it will appear on the bottom of every slide, may be shorthand to save space
{
Ingeniería de Software II - 1er Cuatrimestre 2016 \\ % Your institution for the title page
\medskip
\textit{} % Your email address
}
\date{\today} % Date, can be changed to a custom date

\begin{document}

\begin{frame}
\titlepage % Print the title page as the first slide
\end{frame}

\begin{frame}
\frametitle{Qué vamos a ver} % Table of contents slide, comment this block out to remove it
\tableofcontents % Throughout your presentation, if you choose to use \section{} and \subsection{} commands, these will automatically be printed on this slide as an overview of your presentation
\end{frame}

%----------------------------------------------------------------------------------------
%	PRESENTATION SLIDES
%----------------------------------------------------------------------------------------

%------------------------------------------------
\section{¿Silver Bullet?}
\begin{frame}
\frametitle{¿Silver Bullet?}

\begin{itemize}

\item Un proyecto de software es como un hombre-lobo: usualmente inocentes pero con la capacidad de convertirse en un monstruo.\\
\item Buscamos una bala de plata que disminuya drásticamente los costos de producir software.\\
\item Sin embargo, por la naturaleza misma del software, no hay ningún avance en la tecnología ni en las técnicas de gestión que por sí mismo prometa un aumento en productividad, confiabilidad y simplicidad de siquiera un orden de magnitud de la misma manera que ocurrió con los costos de hardware.\\
\end{itemize}

\end{frame}

\section{Dificultades esenciales vs. Dificultades accidentales}
\begin{frame}
\frametitle{Dificultades esenciales vs. Dificultades accidentales}

Dos tipos de dificultades, vistas a través de un prisma aristotélico:\\
	\begin{itemize}
		\item \textbf{Esenciales}: inherentes a la naturaleza del software.
		\item \textbf{Accidentales}: que tienen que ver con la producción del software y que no son inherentes.

	\end{itemize}
\end{frame}

\section{Esencia}
\begin{frame}
\frametitle{Esencia}

	\begin{itemize}
		\item \textbf{Construcción de conceptos entreladazos}: conjuntos de datos, relaciones entre elementos de datos, algoritmos e invocaciones de funciones.
		\item \textbf{Abstracta}: la contrucción conceptual es la misma bajo muchas diferentes representaciones.				
	\end{itemize}

\textbf{La parte difícil de contruir software es la especificación, diseño y testeo de la construcción conceptual, no representar y probar la fidelidad de la representación.}
\end{frame}

\subsection{Propiedades inherentes a la esencia del software}
\frametitle{Propiedades inherentes a la esencia del software}
\begin{frame}
\frametitle{Propiedades inherentes a la esencia del software (1 / 2)}
\begin{itemize}
	\item \textbf{Complejidad}:
			\begin{itemize}
				\item Las entidades de software son más complejas en tamaño que cualquier otra construcción humana.
				\item Los sistemas tienen gran número de estados lo que los hace difícil de concebir, describir y probar.
				\item Escalar software no es repetir elementos con tamaño más grande: es necesario incrementar la cantidad de elementos.
			\end{itemize}
	\item \textbf{Conformidad}:
			\begin{itemize}
				\item Mucha de la complejidad es arbitraria y viene de la mano de las interfaces con las que un sistema debe conformar.			
			\end{itemize}

\end{itemize}

\end{frame}

\begin{frame}
\frametitle{Propiedades inherentes a la esencia del software (2 / 2)}
\begin{itemize}
	\item \textbf{Modificabilidad}:
			\begin{itemize}
				\item El software de un sistema encarga su función, que es la parte que más siente el peso del cambio, en parte porque es extremadamente maleable.
				\item A veces el cambio viene porque surgen nuevos casos que no pertenecían al dominio original.
				\item Otras, porque el software sobrevive el ciclo de vida de la máquina para la cual fue escrito.			
			\end{itemize}
	\item \textbf{Invisibilidad}:
			\begin{itemize}
				\item El software es invisible e invisualizable.
				\item No tiene una representación espacial.
				\item Diagramar una estructura de software involucra diversos diargramas que miran cosas distintas.
				\item Priva a la mente de usar algunas de sus herramientas conceptuales más potentes y dificulta el diseño y la comunicación. 			
			\end{itemize}

\end{itemize}

\end{frame}

%------------------------------------------------

\subsection{Subsection Example} % A subsection can be created just before a set of slides with a common theme to further break down your presentation into chunks

\begin{frame}
\frametitle{Paragraphs of Text}
Sed iaculis dapibus gravida. Morbi sed tortor erat, nec interdum arcu. Sed id lorem lectus. Quisque viverra augue id sem ornare non aliquam nibh tristique. Aenean in ligula nisl. Nulla sed tellus ipsum. Donec vestibulum ligula non lorem vulputate fermentum accumsan neque mollis.\\~\\

Sed diam enim, sagittis nec condimentum sit amet, ullamcorper sit amet libero. Aliquam vel dui orci, a porta odio. Nullam id suscipit ipsum. Aenean lobortis commodo sem, ut commodo leo gravida vitae. Pellentesque vehicula ante iaculis arcu pretium rutrum eget sit amet purus. Integer ornare nulla quis neque ultrices lobortis. Vestibulum ultrices tincidunt libero, quis commodo erat ullamcorper id.
\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{Bullet Points}
\begin{itemize}
\item Lorem ipsum dolor sit amet, consectetur adipiscing elit
\item Aliquam blandit faucibus nisi, sit amet dapibus enim tempus eu
\item Nulla commodo, erat quis gravida posuere, elit lacus lobortis est, quis porttitor odio mauris at libero
\item Nam cursus est eget velit posuere pellentesque
\item Vestibulum faucibus velit a augue condimentum quis convallis nulla gravida
\end{itemize}
\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{Blocks of Highlighted Text}
\begin{block}{Block 1}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Integer lectus nisl, ultricies in feugiat rutrum, porttitor sit amet augue. Aliquam ut tortor mauris. Sed volutpat ante purus, quis accumsan dolor.
\end{block}

\begin{block}{Block 2}
Pellentesque sed tellus purus. Class aptent taciti sociosqu ad litora torquent per conubia nostra, per inceptos himenaeos. Vestibulum quis magna at risus dictum tempor eu vitae velit.
\end{block}

\begin{block}{Block 3}
Suspendisse tincidunt sagittis gravida. Curabitur condimentum, enim sed venenatis rutrum, ipsum neque consectetur orci, sed blandit justo nisi ac lacus.
\end{block}
\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{Multiple Columns}
\begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

\column{.45\textwidth} % Left column and width
\textbf{Heading}
\begin{enumerate}
\item Statement
\item Explanation
\item Example
\end{enumerate}

\column{.5\textwidth} % Right column and width
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Integer lectus nisl, ultricies in feugiat rutrum, porttitor sit amet augue. Aliquam ut tortor mauris. Sed volutpat ante purus, quis accumsan dolor.

\end{columns}
\end{frame}

%------------------------------------------------
\section{Second Section}
%------------------------------------------------

\begin{frame}
\frametitle{Table}
\begin{table}
\begin{tabular}{l l l}
\toprule
\textbf{Treatments} & \textbf{Response 1} & \textbf{Response 2}\\
\midrule
Treatment 1 & 0.0003262 & 0.562 \\
Treatment 2 & 0.0015681 & 0.910 \\
Treatment 3 & 0.0009271 & 0.296 \\
\bottomrule
\end{tabular}
\caption{Table caption}
\end{table}
\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{Theorem}
\begin{theorem}[Mass--energy equivalence]
$E = mc^2$
\end{theorem}
\end{frame}

%------------------------------------------------

\begin{frame}[fragile] % Need to use the fragile option when verbatim is used in the slide
\frametitle{Verbatim}
\begin{example}[Theorem Slide Code]
\begin{verbatim}
\begin{frame}
\frametitle{Theorem}
\begin{theorem}[Mass--energy equivalence]
$E = mc^2$
\end{theorem}
\end{frame}\end{verbatim}
\end{example}
\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{Figure}
Uncomment the code on this slide to include your own image from the same directory as the template .TeX file.
%\begin{figure}
%\includegraphics[width=0.8\linewidth]{test}
%\end{figure}
\end{frame}

%------------------------------------------------

\begin{frame}[fragile] % Need to use the fragile option when verbatim is used in the slide
\frametitle{Citation}
An example of the \verb|\cite| command to cite within the presentation:\\~

This statement requires citation \cite{p1}.
\end{frame}

%------------------------------------------------

\begin{frame}
\frametitle{References}
\footnotesize{
\begin{thebibliography}{99} % Beamer does not support BibTeX so references must be inserted manually as below
\bibitem[Smith, 2012]{p1} John Smith (2012)
\newblock Title of the publication
\newblock \emph{Journal Name} 12(3), 45 -- 678.
\end{thebibliography}
}
\end{frame}

%------------------------------------------------

\begin{frame}
\Huge{\centerline{The End}}
\end{frame}

%----------------------------------------------------------------------------------------

\end{document}