### 2016-07-11 tp1/informe/referencias.tex

\begin{thebibliography}{9}

\bibitem{Gamma}
  Erich Gamma, Richard Helm, Ralph Johnson, John Vlissides.
  \emph{Design Patterns – Elements of Reusable Object-Oriented Software}.  
  Addison Wesley, 1995.

\bibitem{DoubleDispatch}
    Hebel, K., Johnson, R.. 
    \emph{Arithmetic and double dispatching in Smalltalk-80}.
    University of Illinois at Urbana-Champaign.

\bibitem{Teoricas}
    Santiago Ceria, Hernán Wilkinson, 
    \emph{Clases teóricas de Ingeniería de Software 2}.
    FCEN-UBA,
    2016.

\end{thebibliography}